\documentclass[11pt,a4paper]{scrartcl}
\usepackage[margin=1in,top=1.1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{colortbl}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{scrlayer-scrpage}
% listings uses \rm internally; scrartcl marks it deprecated -- non-fatal
\usepackage[colorlinks=true,linkcolor=mainblue,urlcolor=accentblue,
            citecolor=mainblue,pdfborder={0 0 0}]{hyperref}

% ============================================================
%  Color palette
% ============================================================
\definecolor{mainblue}   {RGB}{  0,  74, 153}   % deep blue
\definecolor{accentblue} {RGB}{ 30, 120, 210}   % medium blue
\definecolor{lightblue}  {RGB}{210, 228, 255}   % pale blue (table heads)
\definecolor{codebg}     {RGB}{245, 248, 252}   % code block background
\definecolor{codeframe}  {RGB}{ 30, 120, 210}   % code left-bar
\definecolor{notebg}     {RGB}{255, 250, 225}   % note box background
\definecolor{notebar}    {RGB}{200, 130,   0}   % note box left bar (amber)
\definecolor{tipbg}      {RGB}{228, 243, 255}   % tip box background
\definecolor{tipbar}     {RGB}{ 30, 120, 210}   % tip box left bar
\definecolor{kwblue}     {RGB}{  0,  90, 190}   % keyword in code
\definecolor{cmtgreen}   {RGB}{ 40, 130,  40}   % comment in code
\definecolor{strred}     {RGB}{170,  20,  20}   % string in code
\definecolor{ruleblue}   {RGB}{  0,  74, 153}   % section rule

% ============================================================
%  KOMA-Script section fonts
% ============================================================
\setkomafont{section}      {\color{mainblue}\large\bfseries}
\setkomafont{subsection}   {\color{mainblue}\normalsize\bfseries}
\setkomafont{subsubsection}{\color{accentblue}\normalsize\bfseries}

% ============================================================
%  Headers and footers
% ============================================================
\pagestyle{scrheadings}
\clearpairofpagestyles
\ohead{\small\textcolor{mainblue!60!black}{VASP Workflow --- User Manual}}
\ihead{\small\textcolor{mainblue!60!black}{\headmark}}
\cfoot{\pagemark}
\addtokomafont{pageheadfoot}{\normalfont\small}

% ============================================================
%  Code listings
% ============================================================
\lstset{
  basicstyle    = \ttfamily\small,
  breaklines    = true,
  breakatwhitespace = true,
  backgroundcolor = \color{codebg},
  frame         = leftline,
  framerule     = 3pt,
  rulecolor     = \color{codeframe},
  framesep      = 8pt,
  xleftmargin   = 14pt,
  keywordstyle  = \color{kwblue}\bfseries,
  commentstyle  = \color{cmtgreen}\itshape,
  stringstyle   = \color{strred},
  upquote       = true,
  numbers       = none,
  aboveskip     = 8pt,
  belowskip     = 4pt,
}

% ============================================================
%  Callout boxes (Note / Tip) using tabular + colortbl
% ============================================================
\newlength{\boxwidth}

\newenvironment{notebox}{%
  \par\vspace{6pt}%
  \setlength{\tabcolsep}{5pt}%
  \setlength{\boxwidth}{\dimexpr\linewidth - 4pt - 2\tabcolsep}%
  \noindent\begin{tabular}{!{\color{notebar}\vrule width 4pt}%
      >{\columncolor{notebg}}p{\boxwidth}}%
  \rule{0pt}{11pt}\textbf{\textcolor{notebar}{Note.}}\enspace\ignorespaces%
}{%
  \rule[-4pt]{0pt}{4pt}%
  \end{tabular}\par\vspace{6pt}%
}

\newenvironment{tipbox}{%
  \par\vspace{6pt}%
  \setlength{\tabcolsep}{5pt}%
  \setlength{\boxwidth}{\dimexpr\linewidth - 4pt - 2\tabcolsep}%
  \noindent\begin{tabular}{!{\color{tipbar}\vrule width 4pt}%
      >{\columncolor{tipbg}}p{\boxwidth}}%
  \rule{0pt}{11pt}\textbf{\textcolor{tipbar}{Tip.}}\enspace\ignorespaces%
}{%
  \rule[-4pt]{0pt}{4pt}%
  \end{tabular}\par\vspace{6pt}%
}

% ============================================================
%  Colored table-header helper
%  Usage: put \thead in the first row of a tabular/tabularx
% ============================================================
\newcommand{\thead}{\rowcolor{lightblue}}

% ============================================================
%  Meta
% ============================================================
\title{%
  \vspace*{-10pt}%
  {\color{mainblue}\rule{\linewidth}{2pt}}\\[8pt]%
  \textbf{\Huge VASP Workflow}\\[4pt]%
  \large User Manual\\[6pt]%
  \normalsize Browser GUI $\cdot$ CLI Agents $\cdot$ SLURM HPC\\[6pt]%
  {\color{mainblue}\rule{\linewidth}{0.5pt}}%
}
\author{Keivan Esfarjani}
\date{v1.9 $\cdot$ July 2026 \quad\textcolor{mainblue}{$|$}\quad
      \url{https://github.com/KeivanS/VASP-Flow}}

% ============================================================
\begin{document}
% ============================================================

\maketitle
\tableofcontents
\newpage

% ============================================================
\section{Introduction}
% ============================================================

VASP Workflow is a framework for setting up, running, and analysing DFT
calculations with VASP.  Two interfaces are provided:

\begin{itemize}
  \item \textbf{Browser GUI} (\texttt{vasp-gui.py}) --- a Flask web
        application: configure calculations through a form, run steps from
        the browser, view band-structure and DOS plots in real time.
  \item \textbf{CLI agents} (\texttt{vasp-agent.py} for workstations,
        \texttt{vasp-agent-slurm.py} for SLURM clusters) --- read a
        plain-text \texttt{instructions.txt} file and generate all INCAR,
        KPOINTS, POSCAR, POTCAR, and job scripts automatically.
\end{itemize}

Both interfaces share the same \texttt{modules/} back-end.

\subsection{Key features}

\begin{itemize}
  \item Functionals: PBEsol, PBE, R2SCAN, HSE06, RVV10, LDA, AM05
  \item Spin-orbit coupling (SOC) and collinear spin (ISPIN=2)
  \item Magnetic materials: correct MAGMOM per atom; spin-resolved band
        and DOS plots (spin-up / spin-down)
  \item GGA+U (Dudarev, LDAUTYPE=2) per element and orbital
  \item Automatic k-path detection via spglib
        (Setyawan--Curtarolo standard paths), with a geometric fallback
  \item Convergence testing: ENCUT and k-mesh sweeps
  \item DOS with orbital projections via \texttt{sumo}
  \item Phonons via phonopy, Wannier90 interface, DFPT Born charges
  \item LOBSTER bonding analysis (COHP/COBI/COOP) via the
        \texttt{08\_lobster} step
  \item Smart KPAR/NCORE auto-defaults scaled with MPI task count
  \item SLURM workflow with dependency-chained job submission
\end{itemize}

\subsection{Requirements}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Package & Notes \\ \midrule
VASP 5.4+ & \texttt{vasp\_std}, \texttt{vasp\_ncl} (SOC), optionally
            \texttt{vasp\_gam} \\
POTCAR library & PAW pseudopotentials in per-element folders \\
Python 3.8+ & Conda environment recommended \\
Flask & \texttt{pip install flask} --- GUI only \\
sumo $\geq$ 2.0 & \texttt{pip install sumo} --- band and DOS plots \\
MPI & \texttt{mpirun}, \texttt{mpiexec}, or \texttt{srun} \\
phonopy & optional; \texttt{conda install -c conda-forge phonopy} \\
wannier90.x & optional; \texttt{conda install -c conda-forge wannier90} \\
\bottomrule
\end{tabularx}

% ============================================================
\section{Installation}
% ============================================================

\begin{lstlisting}[language=bash]
# 1. Clone the repository
git clone https://github.com/KeivanS/VASP-Flow.git
cd VASP-Flow

# 2. Install Python dependencies
pip install flask sumo

# 3. Optional: phonons and Wannier support
conda install -c conda-forge phonopy wannier90

# 4. First-time configuration (creates site.env from template)
make setup
nano site.env     # set VASP paths, POTCAR directory, MPI command
\end{lstlisting}

\subsection{site.env --- platform configuration}

\begin{lstlisting}
PYTHON          = python3
VASP_STD        = /path/to/vasp_std
VASP_NCL        = /path/to/vasp_ncl
VASP_GAM        = /path/to/vasp_gam       # optional
VASP_POTCAR_DIR = /path/to/potpaw_PBE
MPI_LAUNCH      = mpirun -np              # or: srun
MPI_NP          = 16
WANNIER90_X     = wannier90.x
\end{lstlisting}

% ============================================================
\section{Browser GUI}
% ============================================================

\subsection{Starting the GUI}

\begin{lstlisting}[language=bash]
make run          # launches on http://localhost:5001
\end{lstlisting}

\textbf{On a SLURM cluster} --- run on the login node, open on your
laptop via SSH port-forwarding:

\begin{lstlisting}[language=bash]
# On your laptop:
ssh -L 5001:localhost:5001 username@cluster

# On the cluster login node:
make run

# Open http://localhost:5001 in your laptop's browser
\end{lstlisting}

\subsection{Three-tab workflow}

\begin{enumerate}
  \item \textbf{Setup} --- configure the project, then click
        \emph{Generate Workflow}.
  \item \textbf{Workflow} --- run each calculation step; live VASP
        output streams to the browser.
  \item \textbf{Results} --- band-structure and DOS plots; OUTCAR analysis.
\end{enumerate}

To return to a previous project use the \emph{Select project} dropdown.
\textbf{$\to$ Workflow} jumps to the run tab; \textbf{Edit \&
Regenerate} re-opens Setup with all settings pre-filled.

\subsection{Setup tab --- field guide}

\subsubsection{Structure and POTCAR}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Field & Description \\ \midrule
Project name & Folder created in the working directory. \\
POSCAR & Paste the crystal structure, or click \emph{Load file}. \\
POTCAR directory & Path to your PAW library.
  If an element has multiple variants (\texttt{Ga} vs \texttt{Ga\_d})
  a chooser appears. \\
\bottomrule
\end{tabularx}

\subsubsection{Functional}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Choice & Description \\ \midrule
PBEsol & Optimised for solids; recommended starting point. \\
PBE    & Standard GGA; widely used. \\
R2SCAN & Meta-GGA; good accuracy/cost balance. \\
HSE06  & Hybrid --- accurate band gaps, significantly slower. \\
RVV10  & van der Waals; for layered or molecular materials. \\
LDA / AM05 & Rarely needed. \\
\bottomrule
\end{tabularx}

\subsubsection{Spin / SOC}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Choice & Effect \\ \midrule
None & Non-magnetic. \\
Collinear spin (ISPIN=2) & Sets \texttt{ISPIN = 2}; reveals the
  \emph{Magnetic moment / atom} field (Section~\ref{sec:magnetic}). \\
SOC $\parallel$ z/x/y & Sets \texttt{LSORBIT}, \texttt{LNONCOLLINEAR},
  \texttt{SAXIS}; uses the \texttt{vasp\_ncl} binary. \\
\bottomrule
\end{tabularx}

\subsubsection{Other flags}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Field & Effect \\ \midrule
Hexagonal BZ & Correct $M$ point for hexagonal structures. \\
2D / slab & Sets $k_z = 1$ (Gamma-only out-of-plane). \\
GGA+U & Adds Hubbard $U$ rows per element/orbital
  (Section~\ref{sec:ggau}). \\
Execution profile & Select \emph{SLURM HPC} to generate \texttt{sbatch}
  scripts; edit \texttt{profiles/slurm.json} first. \\
MPI tasks & CPU cores per calculation. \\
\bottomrule
\end{tabularx}

\subsection{Task checkboxes}

\begin{tabularx}{\linewidth}{@{}llX@{}}
\thead
Task & Directory & Notes \\ \midrule
Structure relaxation & \texttt{01\_relax} &
  IBRION=2; run first if geometry is not pre-optimised. \\
SCF & \texttt{02\_scf} &
  Required before bands or DOS. \\
Band structure & \texttt{03\_bands} &
  Line-mode KPOINTS; Section~\ref{sec:kpath}. \\
DOS & \texttt{04\_dos} &
  Dense k-mesh; orbital projections via sumo. \\
DFPT & \texttt{06\_dfpt} &
  Born charges, dielectric + piezoelectric tensors;
  Section~\ref{sec:dfpt}. \\
Phonons & \texttt{07\_phonons} &
  Phonopy finite-displacement; DFPT needed for NAC. \\
Wannier90 & \texttt{05\_wannier} &
  NSCF interface; Section~\ref{sec:wannier}. \\
LOBSTER & \texttt{08\_lobster} &
  COHP/COBI/COOP bonding analysis; needs SCF;
  Section~\ref{sec:lobster}. \\
\bottomrule
\end{tabularx}

\begin{notebox}
Normal execution order: \textbf{relaxation $\to$ SCF $\to$ bands $\to$
DOS}.  Each step copies POSCAR/CHGCAR from the previous one automatically.
\end{notebox}

\subsection{Band structure options}

Leave the \emph{k-path} field blank to trigger \textbf{automatic
detection} from the POSCAR lattice vectors (Section~\ref{sec:kpath}).
Enter a path manually to override, e.g.\ \texttt{G-M-K-G}.

\subsection{DOS options}

Add projection rows to request orbital-resolved DOS:

\medskip
\noindent\hspace{1cm}Element: \texttt{Mo}\quad Orbitals: \textbf{d}
\hspace{1.5cm}
Element: \texttt{S}\quad Orbitals: \textbf{p}
\medskip

For spin-polarised calculations sumo plots spin-up positive and
spin-down mirrored below zero automatically.

\subsection{Workflow tab}

Each step shows three controls: \textbf{Run} (submit the step),
\textbf{Edit files} (open INCAR/KPOINTS/POSCAR in the browser; a
\texttt{.bak} is created on first edit), and a \textbf{status badge}.
\textbf{Run All Steps} chains every step and stops on failure.

\subsection{Results tab}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Panel & Contents \\ \midrule
Band structure & Spin-polarised: blue = $\uparrow$, red dashed = $\downarrow$,
  same k-axis, $E_F = 0$.  Non-magnetic: single-colour sumo plot.
  If sumo fails (e.g.\ a pymatgen gap-detection crash) the GUI falls back to a
  built-in \texttt{vasprun.xml} plotter so a band plot still renders. \\
Cumulative DOS & Stacked filled-area by element.
  Spin-polarised: up positive, down mirrored. \\
Projected DOS & Orbital-resolved via sumo; spin-down negated
  automatically. \\
OUTCAR analysis & Energy, Fermi level, magnetic moment, forces. \\
\bottomrule
\end{tabularx}

% ============================================================
\section{CLI Agents}
% ============================================================

\subsection{vasp-agent.py --- workstation}

\begin{lstlisting}[language=bash]
./vasp-agent.py
./vasp-agent.py -i my_inst.txt -s struct.vasp

cd ProjectName
./run_all.sh          # runs every step in sequence
./analyze.sh          # plots after all steps complete
\end{lstlisting}

\subsection{vasp-agent-slurm.py --- SLURM clusters}

\begin{lstlisting}[language=bash]
./vasp-agent-slurm.py                  # uses profiles/slurm.json
./vasp-agent-slurm.py --profile slurm

cd ProjectName
./submit_all.sh        # sbatch with dependency chaining
squeue -u $USER        # monitor jobs
./analyze.sh           # run locally after jobs finish
\end{lstlisting}

Two-phase workflow when convergence tests are requested:

\begin{lstlisting}[language=bash]
./submit_convergence.sh     # STEP 1 -- submit convergence jobs
# review results, then:
./submit_calculations.sh    # STEP 2 -- prompts ENCUT/k-mesh, submits
\end{lstlisting}

\subsection{Generated directory structure}

\begin{lstlisting}
ProjectName/
  instructions.txt
  POSCAR
  POTCAR               built once from $VASP_POTCAR_DIR
  01_relax/            INCAR  KPOINTS  POSCAR  POTCAR  run.sh
  02_scf/
  03_bands/
  04_dos/
  08_lobster/          if LOBSTER task selected (ISYM=0 NSCF + lobster)
  analysis/            plots generated by analyze.sh
  run_all.sh           workstation sequential runner
  submit_all.sh        SLURM job submission (dependency chained)
  analyze.sh           post-processing: plots + OUTCAR extraction
\end{lstlisting}

% ============================================================
\section{High-Throughput Screening (Materials Project)}
\label{sec:highthroughput}
% ============================================================

\texttt{ht-mp-scf.py} drives a batch of SCF + ELF calculations straight
from Materials Project IDs.  For every \texttt{mp-id} it downloads the
\textbf{primitive} cell from MP, stages a per-material
\texttt{instructions.txt} (an SCF task plus an \texttt{INCAR scf:} block
setting \texttt{LELF = .TRUE.} for the electron localization function),
and writes a \texttt{runall.sh} that calls the chosen agent to build the
inputs and then runs or submits each job.  Each SCF uses the automatic
\texttt{coarse} density (1000~kpra, cell-aware) by default; override via
\texttt{KMESH\_DENSITY: fine} in \texttt{\_ht\_inputs/<id>/instructions.txt}
or supply an explicit \texttt{-{}-kmesh} grid.

\begin{lstlisting}[language=bash]
pip install pymatgen mp-api          # MP download + primitive standardisation
export MP_API_KEY=...                 # Materials Project API key
export VASP_POTCAR_DIR=...            # POTCAR library (used by the agent)

# one mp-id per line (blank lines and # comments ignored)
printf 'mp-149\nmp-2534\n' > highthrouput_list

./ht-mp-scf.py --agent local  --mpi 16        # workstation
./ht-mp-scf.py --agent slurm  --profile slurm # cluster (chained by default)

./runall.sh                           # build inputs + run/submit every job
\end{lstlisting}

The ELFCAR for each material is written to \texttt{<mp-id>/02\_scf/}.
A summary \texttt{\_ht\_inputs/elements\_Z.txt} tabulates every element in
the screening set with its atomic number $Z$ (from a built-in table --- no
MP query), plus a per-material breakdown.

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Option & Meaning \\ \midrule
\texttt{-{}-agent} & \texttt{local} (vasp-agent.py) or \texttt{slurm}
  (vasp-agent-slurm.py); prompts if omitted. \\
\texttt{-{}-list} & ID file; default \texttt{highthrouput\_list}. \\
\texttt{-{}-functional} & PBE (default), PBEsol, R2SCAN, \ldots \\
\texttt{-{}-mpi} & MPI tasks per job. \\
\texttt{-{}-encut} & Optional ENCUT override (eV). \\
\texttt{-{}-kmesh} & Explicit SCF $\Gamma$-mesh, e.g.\ \texttt{'8 8 8'} or
  \texttt{'12'} (cubic); overrides the automatic density.  Without this flag,
  the cell-aware coarse (1000~kpra) density is used.  Per-material overrides
  via the \texttt{KMESH} or \texttt{KMESH\_DENSITY} line in
  \texttt{instructions.txt}. \\
\texttt{-{}-chain} / \texttt{-{}-no-chain} & SLURM only; see below. \\
\bottomrule
\end{tabularx}

\subsection{Execution order}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Mode & Behaviour \\ \midrule
local & \texttt{run\_all.sh} blocks --- materials run strictly
  one-at-a-time. \\
slurm \texttt{-{}-chain} (default) & Each material's SCF is submitted with
  \texttt{-{}-dependency=afterok} on the previous one, so the cluster runs
  them sequentially even though all are queued up front. \\
slurm \texttt{-{}-no-chain} & Materials are submitted independently and run
  whenever the scheduler allows. \\
\bottomrule
\end{tabularx}

\begin{notebox}
The driver fetches structures over the network (Materials Project), so
run it on a node with internet --- typically the login node.
\texttt{runall.sh} performs no downloads and can run on a compute node.
\end{notebox}

% ============================================================
\section{instructions.txt Format}
\label{sec:instructions}
% ============================================================

\subsection{Example}

\begin{lstlisting}
Project: My material name

Methods: PBEsol functional
         collinear magnetization
         GGA+U with U=4.0 on Fe-d orbitals

Tasks: structure relaxation
       SCF calculation
       band structure
       DOS (projected on Fe-d and O-p)

Convergence: Test k-points 6x6x6 to 12x12x12
             ENCUT 400-600 eV

MAGMOM: 4.0
NSW: 100
EDIFFG: -0.01
MPI: 128
NODES: 2
NTASKS_PER_NODE: 64
WALLTIME: 12:00:00
\end{lstlisting}

The parser is case-insensitive and searches the entire file, so section
order is flexible.

\subsection{Keyword reference}
\label{sec:keywords}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Keyword & Meaning \\ \midrule
\texttt{Project:} & Project name; becomes the output folder. \\
\texttt{MAGMOM:} & Initial moment in $\mu_{\mathrm{B}}$/atom (collinear only).
  Written as \texttt{MAGMOM = N*value}. Default 2.0. \\
\texttt{ISIF:} & Relaxation type: 3=full, 2=ions only, 4=fixed volume. \\
\texttt{NSW:} & Max ionic steps. Default 100. \\
\texttt{EDIFFG:} & Force criterion in eV/\AA. Default $-0.01$. \\
\texttt{ENCUT:} & Manual plane-wave cutoff in eV. \\
\texttt{KMESH:} & Explicit Gamma k-mesh overriding the automatic density,
  e.g.\ \texttt{8 8 8} or \texttt{12} (cubic).  Applies to every automatic
  mesh (relax/SCF/DOS); for 2-D give the third value (\texttt{12 12 1}). \\
\texttt{KMESH\_DENSITY:} & Automatic k-mesh density tier: \texttt{coarse}
  (1000~kpra, default for relaxation and convergence tests) or \texttt{fine}
  (5000~kpra, default for SCF/DOS/spectroscopic steps).  kpra $= N_k \times
  N_\mathrm{atoms}$ where $N_k = N_x N_y N_z$; subdivisions scale as
  $N_i \propto |\mathbf{b}_i^*|$.  Overridden by explicit \texttt{KMESH}. \\
\texttt{PRESSURE:} & External pressure for constant-$P$ relaxation, e.g.\
  \texttt{10 GPa} or \texttt{50 kbar}.  Forces \texttt{ISIF=3},
  \texttt{IBRION=2}, \texttt{PSTRESS} (Section~\ref{sec:pressure}). \\
\texttt{INCAR: \ldots\ END\_INCAR} & Block of literal INCAR tags merged
  into the generated INCAR(s); optionally per-step
  (Section~\ref{sec:incar-raw}). \\
\texttt{NKPTS:} & k-points per band segment. Default 40. \\
\texttt{MPI:} & Total MPI tasks (nodes $\times$ cores/node). \\
\texttt{KPAR:} & Global k-point parallelisation. \\
\texttt{NCORE:} & Global cores-per-orbital. \\
\texttt{SCF\_KPAR:} \textit{etc.} &
  Per-step KPAR: also BANDS\_, DOS\_, RELAX\_, PHONONS\_
  (DFPT is forced to KPAR${}={}$NCORE${}={}$1). \\
\texttt{SCF\_NCORE:} \textit{etc.} & Per-step NCORE; same prefixes. \\
\texttt{NODES:} & Compute nodes (SLURM). \\
\texttt{NTASKS\_PER\_NODE:} & MPI tasks per node (SLURM). \\
\texttt{PARTITION:} & SLURM partition. \\
\texttt{WALLTIME:} & Time limit, e.g.\ \texttt{12:00:00}. \\
\texttt{ACCOUNT:} & SLURM account / allocation. \\
\texttt{DFPT\_EDIFF:} & EDIFF for DFPT. Default \texttt{1E-8}. \\
\texttt{PHONONS\_DIM:} & Supercell, e.g.\ \texttt{2 2 2}. \\
\texttt{PHONONS\_MESH:} & q-mesh, e.g.\ \texttt{20 20 20}. \\
\texttt{PHONONS\_DISP:} & Displacement in \AA. Default 0.01. \\
\texttt{PHONONS\_BAND:} & High-symmetry path for phonon bands. \\
\texttt{PHONONS\_NAC:} & \texttt{FALSE} to disable NAC. \\
\texttt{WANNIER\_NUM\_WANN:} & Number of Wannier functions. \\
\texttt{WANNIER\_PROJ:} & Projections, e.g.\ \texttt{Ga:s;As:p}. \\
\texttt{WANNIER\_EWIN:} & Energy window, e.g.\ \texttt{-5.0 to 15.0}. \\
\bottomrule
\end{tabularx}

% ============================================================
\section{Methods}
% ============================================================

\subsection{Functionals}

\begin{lstlisting}
Methods: PBEsol functional    # recommended for solids
Methods: R2SCAN functional    # meta-GGA
Methods: HSE06 functional     # hybrid -- more accurate but slower
Methods: RVV10 functional     # van der Waals
\end{lstlisting}

Default ENCUT values: PBE 400~eV, PBEsol 450~eV, R2SCAN 680~eV,
HSE06 400~eV, RVV10 450~eV.

\subsection{Spin-orbit coupling}

\begin{lstlisting}
Methods: SOC with magnetization in z-direction
\end{lstlisting}

Uses \texttt{vasp\_ncl}.  Sets \texttt{LSORBIT}, \texttt{LNONCOLLINEAR},
and \texttt{SAXIS}.

\subsection{Magnetic materials (collinear spin)}
\label{sec:magnetic}

\begin{lstlisting}
Methods: collinear magnetization
MAGMOM: 4.0
\end{lstlisting}

Sets \texttt{ISPIN = 2} and \texttt{MAGMOM = N * value} in every INCAR,
where $N$ is the atom count from the POSCAR.

\begin{tipbox}
Typical starting moments: Fe, Co $\approx$ 4--5~$\mu_{\mathrm{B}}$;
Ni $\approx$ 2~$\mu_{\mathrm{B}}$; Cr, Mn $\approx$ 3--4~$\mu_{\mathrm{B}}$;
rare earths $\approx$ 7~$\mu_{\mathrm{B}}$.
VASP converges the self-consistent value automatically.
\end{tipbox}

\textbf{Antiferromagnets} --- after generation, edit INCAR to alternate
signs:
\begin{lstlisting}
MAGMOM = 4.0 -4.0 4.0 -4.0    ! alternating per sublattice
\end{lstlisting}

\textbf{Spin-resolved plots} produced automatically after DOS/bands run:

\begin{itemize}
  \item \emph{Cumulative DOS} --- spin-up above the x-axis; spin-down
        mirrored below zero; elements colour-coded; solid = $\uparrow$,
        dashed = $\downarrow$.
  \item \emph{Band structure} --- spin-up in blue, spin-down in red
        dashed, same k-axis, $E_F = 0$.
  \item \emph{Projected DOS} --- sumo natively negates the spin-down
        channel in the orbital-resolved PDF.
\end{itemize}

Check the converged moment:
\begin{lstlisting}[language=bash]
grep "magnetization (x)" OUTCAR | tail -5
\end{lstlisting}

\subsection{GGA+U}
\label{sec:ggau}

\begin{lstlisting}
Methods: GGA+U with U=4.0 on Fe-d orbitals
         GGA+U with U=3.0 on Ni-d orbitals
         GGA+U with U=6.0 on Ce-f orbitals
\end{lstlisting}

Dudarev scheme (LDAUTYPE=2); $U_{\mathrm{eff}} = U - J$ with $J = 0$ set
automatically.  In the Dudarev formalism the parameter labelled $U$ is already
$U_{\mathrm{eff}}$, so no separate $J$ entry is needed.
Combine freely with \texttt{collinear magnetization}.

The automatic lookup is triggered whenever the structure contains a
transition-metal or rare-earth element \emph{and} at least one
electronegative anion from the set
O, F, S, Se, Te, Cl, Br, I.
This covers oxides, sulfides (e.g.\ AgCrP$_2$S$_6$), selenides, tellurides,
and halides.  U values are read from \texttt{hubbard\_u\_defaults.csv};
elements absent from the table (e.g.\ Ag with a filled $d^{10}$ shell)
receive no correction.

% ============================================================
\subsection{Automatic k-mesh generation}
\label{sec:kmesh}
% ============================================================

All Gamma-centred meshes are generated automatically from the POSCAR geometry.
Two density tiers are available via the \texttt{KMESH\_DENSITY} key
(Section~\ref{sec:keywords}):

\begin{tabularx}{\linewidth}{@{}lllX@{}}
\thead
Tier & kpra & Typical use & Notes \\ \midrule
\texttt{coarse} & 1000 & Relaxation, convergence tests & Agent k-point sweeps start here \\
\texttt{fine}   & 5000 & SCF, DOS, spectroscopic steps & Default for all production steps \\
\bottomrule
\end{tabularx}

\medskip
\textbf{kpra} (k-points per reciprocal atom) $= N_x N_y N_z \times N_\mathrm{atoms}$.
Given a target kpra and the reciprocal lattice vectors $\mathbf{b}_i^*$
(columns of $2\pi(\mathbf{A}^{-1})^\top$), the algorithm solves

\[
  N_i = \lceil \alpha \, |\mathbf{b}_i^*| \rceil_{\text{even}}, \qquad
  \alpha = \left(\frac{N_k}{|\mathbf{b}_1^*||\mathbf{b}_2^*||\mathbf{b}_3^*|}\right)^{1/3},
  \quad N_k = \mathrm{kpra}/N_\mathrm{atoms}.
\]

Each $N_i$ is rounded to the nearest even integer so that BZ boundary points
($k = 1/2$) are always included.  For 2-D slabs $N_z$ is fixed to 1.

\textbf{Hexagonal cells} ($a = b$, $\gamma = 120°$): in-plane subdivisions
are fixed at 6 (coarse) or 12 (fine); $N_z$ is still computed from
$|\mathbf{b}_3^*|/|\mathbf{b}_1^*|$ to maintain uniform sampling.

\textbf{DOS mesh}: always $2N_x \times 2N_y \times 2N_z$, ensuring denser
sampling for accurate spectral features.  The DFPT mesh is forced to match
the SCF mesh exactly.

An explicit \texttt{KMESH: N1 N2 N3} entry overrides the automatic density
for all steps.

% ============================================================
\section{Advanced INCAR Control}
\label{sec:incar}
% ============================================================

\textbf{Any VASP keyword can be specified explicitly} --- nothing is
hard-wired.  Three routes, in increasing precedence:
(1)~an \texttt{INCAR:} \ldots\ \texttt{END\_INCAR} block in the
instructions file (global or per-step);
(2)~the agents' \texttt{--incar} command-line option;
(3)~direct editing --- in the GUI every generated INCAR is editable per
step (Workflow tab $\to$ \emph{Edit files}, with an automatic
\texttt{.bak} backup), and of course the files on disk can be edited
before running.  This section covers the first two plus
\textbf{constant-pressure relaxation}.

\subsection{Raw INCAR passthrough}
\label{sec:incar-raw}

Place literal INCAR lines --- in ordinary VASP syntax --- inside an
\texttt{INCAR:} \ldots\ \texttt{END\_INCAR} block in your
\texttt{instructions.txt}.  They are merged into the generated INCAR(s):

\begin{lstlisting}
INCAR:                  # global block -- applies to every step
   LREAL  = Auto
   ALGO   = Fast
   NELMIN = 6
END_INCAR
\end{lstlisting}

A tag that the agent already writes is \textbf{overwritten in place}
(so \texttt{ALGO = Fast} above replaces the default \texttt{ALGO =
Normal}); any tag the agent does not normally emit is appended under a
clearly labelled \texttt{\# User INCAR overrides} section, so nothing is
silently lost.

To target a single step, name it on the header line.  Recognised steps
are \texttt{relax}, \texttt{scf}, \texttt{bands}, \texttt{dos},
\texttt{wannier}, \texttt{dfpt}, \texttt{phonons}, \texttt{lobster}:

\begin{lstlisting}
INCAR scf:              # applies only to 02_scf
   EDIFF = 1E-8
   SIGMA = 0.10
END_INCAR
\end{lstlisting}

\begin{notebox}
Precedence: a \emph{per-step} block beats the \emph{global} block, and
both beat the agent's generated defaults.  This is the recommended way to
set tags the parser has no dedicated keyword for (e.g.\ \texttt{LMAXMIX},
\texttt{ADDGRID}, \texttt{LVTOT}, \texttt{IVDW}, \texttt{AMIX}).
Full-line \texttt{\#} comments in the instructions file are ignored by
the parser, so a task keyword inside a comment does not enable that task.
\end{notebox}

\subsection{Command-line INCAR overrides (\texttt{--incar})}
\label{sec:incar-cli}

Both \texttt{vasp-agent.py} and \texttt{vasp-agent-slurm.py} accept a
repeatable \texttt{--incar} option that injects explicit VASP tags at
invocation time, without touching the instructions file:

\begin{lstlisting}[language=bash]
./vasp-agent.py -i instructions.txt -s POSCAR \
    --incar "ALGO=Fast; NELMIN=6" \      # every step
    --incar "scf:NBANDS=64"              # 02_scf only
\end{lstlisting}

Multiple tags in one option are separated by \texttt{;} (not commas ---
values such as \texttt{MAGMOM} may contain commas).  An optional step
prefix (\texttt{relax}/\texttt{scf}/\texttt{bands}/\texttt{dos}/%
\texttt{wannier}/\texttt{dfpt}/\texttt{phonons}/\texttt{lobster})
targets a single step.  Command-line tags are applied \emph{after} the
instructions-file blocks, so they win on conflicts.

\begin{notebox}
Exceptions: physics-critical guards are re-imposed regardless of
overrides --- \texttt{ISYM=2} and \texttt{KPAR=NCORE=1} in
\texttt{06\_dfpt}, and the symmetry-off \texttt{ISYM} in
\texttt{08\_lobster}.
\end{notebox}

\subsection{Constant-pressure relaxation}
\label{sec:pressure}

A single \texttt{PRESSURE} line turns the relaxation into a
constant-pressure optimisation: the agent forces \texttt{ISIF = 3} (cell
shape + volume + ions) and \texttt{IBRION = 2}, and imposes the target
pressure through \texttt{PSTRESS}.

\begin{lstlisting}
PRESSURE: 10 GPa        # also accepted: "5 kbar", "constant pressure 8"
\end{lstlisting}

VASP's \texttt{PSTRESS} tag is in \textbf{kBar} (1~GPa $=$ 10~kBar).  The
agent converts automatically; if no unit is given, \textbf{GPa is
assumed}.  The resulting \texttt{01\_relax/INCAR} reads, e.g.:

\begin{lstlisting}
# Constant-pressure relaxation @ 10 GPa  (PSTRESS = 100 kBar)
IBRION = 2
ISIF   = 3
NSW    = 100
EDIFFG = -0.01
PSTRESS = 100   ! external pressure in kBar (10 GPa)
\end{lstlisting}

\begin{tipbox}
\texttt{PSTRESS} adds the $PV$ term so the relaxed cell sits at the
requested external pressure.  Use a tight \texttt{EDIFFG} and a
well-converged \texttt{ENCUT}; the Pulay stress error grows at high
pressure, so verify with a higher \texttt{ENCUT} before trusting the
equilibrium volume.
\end{tipbox}

% ============================================================
\section{K-Path for Band Structure}
\label{sec:kpath}
% ============================================================

\subsection{Auto-detection (recommended)}

Omit the path in \texttt{instructions.txt} or leave the field blank in
the GUI:

\begin{lstlisting}
Tasks: band structure        # no "along ..." -> auto-detect from POSCAR
\end{lstlisting}

Since v1.6 the generator first determines the \emph{actual space group}
of the POSCAR with spglib (via pymatgen's
\texttt{KPathSetyawanCurtarolo}) and uses the standard
Setyawan--Curtarolo high-symmetry path for that space group.  Every
high-symmetry point is transformed through Cartesian reciprocal space
into the reciprocal basis of \emph{this} POSCAR, so the path is correct
for the actual cell used in the SCF regardless of orientation or
setting.

If spglib or pymatgen is not installed (or the symmetry analysis
fails), the generator falls back to the geometric classifier below: it
reads the POSCAR lattice vectors, computes cell lengths and angles, and
selects a default path by Bravais type.  2-D structures always use the
geometric route so the in-plane path is preserved.

\begin{tabularx}{\linewidth}{@{}llX@{}}
\thead
Type & Detection criterion & Default path \\ \midrule
cF (FCC) & $a{=}b{=}c$, angles $\approx 60^\circ$
         & G-X-W-K-G-L-U-W \\
cI (BCC) & $a{=}b{=}c$, angles $\approx 109.5^\circ$
         & G-H-N-G-P-H \\
cP (cubic) & $a{=}b{=}c$, $90^\circ$
          & G-X-M-G-R-X \\
hP (hexagonal) & $a{\approx}b$, $\gamma{=}120^\circ$
              & G-M-K-G-A-L-H-A \\
tP (tetragonal) & $a{\approx}b{\neq}c$, $90^\circ$
               & G-X-M-G-Z-R-A-Z \\
oP (orthorhombic) & $90^\circ$, $a{\neq}b{\neq}c$
                 & G-X-S-Y-G-Z-U-R-T-Z \\
rP (rhombohedral) & $a{=}b{=}c$, equal angles
                 & G-T-F-G-L \\
mP (monoclinic) & $\alpha{=}\gamma{=}90^\circ$, $\beta{\neq}90^\circ$
               & G-Y-A-Z-G-B-D \\
\bottomrule
\end{tabularx}

\begin{notebox}
For \textbf{conventional} FCC or BCC cells (right angles, 4 or 2 atoms),
auto-detection returns the cubic (cP) path.
Use the \textbf{primitive} cell for an unfolded FCC/BCC band structure ---
the primitive FCC cell has angles $\approx 60^\circ$ and is recognised
automatically.
\end{notebox}

The detection route appears in the KPOINTS comment:
\begin{lstlisting}
Line-mode KPOINTS  [spglib ... -- auto-detected]   # spglib route
Line-mode KPOINTS  [FCC -- auto-detected]          # geometric fallback
\end{lstlisting}

\subsection{Manual override}

\begin{lstlisting}
Tasks: band structure along G-X-W-K-G-L
\end{lstlisting}

\begin{tabularx}{\linewidth}{@{}lllX@{}}
\thead
Label & Coords & Use & \\ \midrule
G & (0, 0, 0) & all & $\Gamma$ \\
M & (0.5, 0.5, 0) & cubic, tet & (0.5, 0, 0) in hexagonal \\
K & (1/3, 1/3, 0) & hexagonal & \\
R & (0.5, 0.5, 0.5) & cubic & \\
Z & (0, 0, 0.5) & general & \\
A & (0, 0, 0.5) & hex, tet & \\
L & (0.5, 0, 0.5) & hexagonal & \\
H & (1/3, 1/3, 0.5) & hexagonal & \\
S & (0.5, 0.5, 0) & orthorhombic & \\
W & (0.5, 0.25, 0.75) & \textbf{FCC primitive} & \\
P & (0.25, 0.25, 0.25) & \textbf{BCC primitive} & \\
N & (0, 0, 0.5) & \textbf{BCC primitive} & \\
\bottomrule
\end{tabularx}

% ============================================================
\section{DOS Projections}
% ============================================================

All of the following forms are accepted:

\begin{lstlisting}
DOS (projected on Mo-d and S-p)
DOS with projections on Ni:d and Ni:s
DOS with projection on Fe-d and O-p
\end{lstlisting}

Orbital channels: \texttt{s}, \texttt{p}, \texttt{d}, \texttt{f}.
Separator is either \texttt{-} or \texttt{:}.
Sumo generates an orbital-resolved PDF in \texttt{analysis/}.

% ============================================================
\section{ELF Bond Profiles}
% ============================================================

If an SCF calculation writes an \texttt{ELFCAR} (electron localization
function, requested with \texttt{LELF = .TRUE.} --- the high-throughput
driver sets this automatically), the utility \texttt{elf\_bonds.py} plots ELF
along each non-equivalent nearest-neighbour bond, sampled from one atom centre
to the other:

\begin{lstlisting}[language=bash]
./elf_bonds.py [ELFCAR] [--out PREFIX] [--nn-scale 1.2] [--npoints 200]
\end{lstlisting}

It parses the \texttt{ELFCAR} grid directly (only numpy + matplotlib), groups
bonds by element pair and length (rounded to 0.01~\AA{}) to pick the
non-equivalent ones, and interpolates ELF along each bond with periodic
boundary conditions.  Output is \texttt{FORMULA\_elf\_bonds.png/pdf} (the prefix
defaults to the reduced formula, e.g.\ \texttt{MoSe2\_elf\_bonds.pdf}; override
with \texttt{--out}) showing ELF vs.\ distance along the bond, with the
ELF~$=0.5$ homogeneous-electron-gas reference marked.  \texttt{analyze.sh}
runs it automatically when \texttt{02\_scf/ELFCAR} is present, and the GUI
Results tab shows the same plot whenever the SCF wrote an \texttt{ELFCAR}.

\begin{tipbox}
Reading the profile: ELF is high (near~1) where electrons are paired and
localized.  A covalent bond therefore shows an ELF \emph{maximum} in the
bonding region \emph{between} the atoms (e.g.\ Si--Si $\approx 0.95$ at
mid-bond), not at the nuclei.  Because VASP builds the \texttt{ELFCAR} from the
pseudo (valence) density, the value \emph{at} an atom centre is a PAW artifact
--- often low, even $\approx 0$ for heavy atoms --- so read the bonding region,
not the endpoints.
\end{tipbox}

\begin{notebox}
VASP does \emph{not} compute ELF for non-collinear / spin-orbit runs
(\texttt{LSORBIT}/\texttt{LNONCOLLINEAR}); it prints
\texttt{"WARNING: ELF not implemented for non collinear case"} and writes an
all-zero \texttt{ELFCAR}.  \texttt{elf\_bonds.py} detects this and reports it
rather than producing a blank plot --- re-run the SCF \emph{without} SOC to get
ELF.
\end{notebox}

% ============================================================
\section{Convergence Tests}
% ============================================================

\begin{lstlisting}
# Range (auto-steps)
Convergence: Test k-points from 6x6x1 to 18x18x1    # 2D material
             Test k-points from 6x6x6 to 12x12x12   # 3D material
             ENCUT 400-600 eV

# Explicit mesh list
Convergence: Test k-points 6x6x3, 12x12x6, 18x18x9
\end{lstlisting}

Results in \texttt{00\_convergence/kpoints/} and
\texttt{00\_convergence/encut/}.

\begin{lstlisting}[language=bash]
./run_convergence.sh          # workstation
./submit_convergence.sh       # SLURM
./analyze_convergence.sh      # plots: energy / pressure / forces
\end{lstlisting}

\subsection{Shifted test structure}

The POSCAR used in \texttt{00\_convergence} has \textbf{atom 1 displaced
by $(0.01,\,0.02,\,0.03)$\,\AA{} (Cartesian)} so that the force on it is
non-zero and its convergence can be tracked --- at the equilibrium
(high-symmetry) positions the forces would vanish identically and the
force panel would be meaningless.  The displacement is converted through
the lattice for Direct-mode POSCARs.  All production steps
(\texttt{01\_relax}, \texttt{02\_scf}, \dots) use the \emph{original,
unshifted} positions.

\subsection{Automatic parameter selection}

After the runs, \texttt{00\_convergence/choose\_params.py} (invoked by
\texttt{run\_convergence.sh}, \texttt{analyze\_convergence.sh}, and the
phase-2 scripts) selects the smallest ENCUT and k-mesh from which every
subsequent change stays below
\[
  \max_i \bigl|\Delta F_i^{\text{atom 1}}\bigr| < 5~\text{meV/\AA}
  \quad\text{and}\quad
  \max_i \bigl|\Delta \sigma_{ii}\bigr| < 0.5~\text{kbar},
\]
and writes the result to \texttt{00\_convergence/converged\_params.txt}.
Phase 2 (\texttt{run\_calculations.sh}, \texttt{submit\_calculations.sh},
and the GUI form) offers these values as defaults; when run
\emph{non-interactively} --- e.g.\ a convergence test automatically
followed by the SCF in a chained/nohup run --- they are applied without
prompting.  If the criteria are not met within the tested range, the
largest tested value is used and a \texttt{NOT CONVERGED} warning is
issued.  Thresholds can be overridden:
\texttt{choose\_params.py --ftol 3 --ptol 0.2}.

% ============================================================
\section{Parallelization}
% ============================================================

\subsection{Auto-defaults}

When KPAR/NCORE are not specified the agent computes from total tasks $N$:

\begin{tabularx}{\linewidth}{@{}llll@{}}
\thead
Step & KPAR & NCORE & Notes \\ \midrule
SCF / DOS / relax & $\lfloor\sqrt{N}\rfloor$ & $N/\text{KPAR}$ & \\
bands & 1 & $N$ & Required for line-mode \\
DFPT & 1 & 1 & Forced: linear response / PEAD supports neither \\
phonons & 1 & $\lfloor\sqrt{N}\rfloor$ & \\
\bottomrule
\end{tabularx}

\begin{notebox}
KPAR $\times$ NCORE must divide $N$ evenly.  The agent rounds down
automatically to the nearest valid value.
\end{notebox}

\subsection{Manual overrides}

\begin{lstlisting}
MPI: 128        KPAR: 4        NCORE: 8

# Per-step (take priority over globals)
SCF_KPAR: 4     BANDS_KPAR: 1     DOS_KPAR: 8
SCF_NCORE: 8    BANDS_NCORE: 16   DOS_NCORE: 8
RELAX_KPAR: 2   PHONONS_KPAR: 1   # DFPT is always KPAR=1, NCORE=1
\end{lstlisting}

% ============================================================
\section{SLURM Cluster Settings}
% ============================================================

\subsection{profiles/slurm.json}

\begin{lstlisting}
{
  "vasp_std": "vasp_std",  "vasp_ncl": "vasp_ncl",
  "mpi_cmd":  "srun",
  "modules":  ["intel/2021", "impi/2021", "vasp/6.3.2"],
  "slurm": {
    "partition":       "standard",
    "nodes":           2,
    "ntasks_per_node": 64,
    "time":            "12:00:00",
    "account":         "mygroup"
  }
}
\end{lstlisting}

\subsection{Per-project overrides (in instructions.txt)}

\begin{lstlisting}
NODES: 4     NTASKS_PER_NODE: 32     PARTITION: gpu
WALLTIME: 48:00:00     ACCOUNT: myallocation
\end{lstlisting}

% ============================================================
\section{Wannier90}
\label{sec:wannier}
% ============================================================

\begin{lstlisting}
Tasks: Wannier90 wannierization

WANNIER_NUM_WANN: 8
WANNIER_PROJ: Ga:s;As:p
WANNIER_EWIN: -5.0 to 15.0
\end{lstlisting}

The agent generates \texttt{05\_wannier/wannier90.win} with the NSCF
k-mesh and projection block.  After the NSCF run, execute
\texttt{wannier90.x} from the \texttt{05\_wannier/} directory.

% ============================================================
\section{Born Charges, Dielectric \& Piezoelectric Tensors (DFPT)}
\label{sec:dfpt}
% ============================================================

The \texttt{DFPT} task creates \texttt{06\_dfpt}, which computes Born
effective charges $Z^{*}$, the macroscopic dielectric tensor, and the
piezoelectric stress tensor from the converged SCF charge density.

Two routes are generated automatically depending on the method:

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Method & Route \\ \midrule
PBE, PBEsol, LDA, AM05 (no SOC) &
  Linear response: \texttt{IBRION=8}, \texttt{LEPSILON=.TRUE.} \\
SOC, HSE06, or R2SCAN &
  Finite field (PEAD): \texttt{IBRION=6}, \texttt{LCALCEPS=.TRUE.}
  --- VASP's DFPT is not implemented for noncollinear magnetism,
  hybrids, or meta-GGAs.  Requires an insulating system. \\
\bottomrule
\end{tabularx}

\begin{notebox}
Both routes support neither k-point nor band parallelisation:
\textbf{KPAR and NCORE are forced to 1} in the generated INCAR
(overriding any global or per-step setting).  The full MPI task count
is still used for plane-wave parallelisation.
Likewise \textbf{ISYM=2 is enforced} (symmetry imposed): perturbation
theory relies on the symmetry-reduced perturbation set, so a
symmetry-off \texttt{ISYM=0}/\texttt{$-$1} --- e.g.\ from a global
\texttt{INCAR:} override block intended for the LOBSTER/NSCF steps ---
is never inherited by \texttt{06\_dfpt}.
\end{notebox}

\texttt{copy\_from\_scf.sh} (invoked by \texttt{run.sh}) pulls
\texttt{WAVECAR} and \texttt{CHGCAR} from \texttt{02\_scf} at run time.
If \texttt{02\_scf} has been converted in place for LOBSTER
(\texttt{ISYM=0}/\texttt{$-$1} in its INCAR, as done by the
high-throughput \texttt{run\_scf\_for\_lobster.sh}), the
\texttt{WAVECAR} copy is skipped --- it holds the full, unsymmetrised
k-mesh, which the ISYM=2 run cannot restart from --- and VASP
regenerates the orbitals from the \texttt{CHGCAR}, which the LOBSTER
conversion (a fixed-charge NSCF with \texttt{LCHARG=.FALSE.}) leaves
untouched.  If the symmetry-off INCAR is \emph{not} a fixed-charge
NSCF, a warning is printed that the \texttt{CHGCAR} itself may stem
from a symmetry-off SCF and a symmetric SCF should be re-run first.

\texttt{run.sh} executes \texttt{extract\_born.py} after VASP, which
writes:

\begin{itemize}
  \item \texttt{BORN} --- phonopy NAC format: electronic
        $\varepsilon_{\infty}$ + $Z^{*}$ per atom.
  \item \texttt{born\_charges.txt} --- human-readable summary shown in
        the GUI Results tab: electronic (ion-clamped)
        $\varepsilon_{\infty}$, ionic contribution, total static
        $\varepsilon_{0}=\varepsilon_{\infty}+\varepsilon_{\mathrm{ion}}$;
        $Z^{*}$ tensors labelled by element; an acoustic-sum-rule check
        ($\sum_{\mathrm{ions}} Z^{*}$ should vanish --- a large residual
        indicates an under-converged k-mesh or EDIFF); and the
        piezoelectric stress tensors $e$ (C/m$^2$, clamped-ion, ionic,
        and relaxed-ion total; meaningful only for non-centrosymmetric
        crystals).
\end{itemize}

\texttt{DFPT\_EDIFF} in \texttt{instructions.txt} tightens the
electronic convergence (default \texttt{1E-8}).

\subsection{Example}

\begin{lstlisting}
Project: GaAs
Methods: PBE functional
Tasks: relaxation, SCF, DFPT Born charges and dielectric
DFPT_EDIFF: 1E-8
MPI: 16
\end{lstlisting}

In the GUI, tick \textbf{Born Charges \& Dielectric Constant (DFPT)} on
the Setup tab (the DFPT EDIFF field is optional); after running
\texttt{06\_dfpt} from the Workflow tab, the full summary appears in
the \textbf{Born charges} panel of the Results tab.  With SOC, HSE06
or R2SCAN selected under Methods, the finite-field INCAR is generated
automatically --- no extra input is needed.  For LO-TO splitting in
phonon spectra, combine with the phonon task as in the GaAs worked
example (Section~\ref{sec:example-gaas-nac}).

% ============================================================
\section{Phonons (phonopy)}
% ============================================================

\begin{lstlisting}
Tasks: phonons lattice dynamics

PHONONS_DIM: 2 2 2
PHONONS_MESH: 20 20 20
PHONONS_DISP: 0.01
PHONONS_BAND: G M K G A
PHONONS_NAC: FALSE      # disable non-analytic correction
\end{lstlisting}

\begin{notebox}
For the non-analytic correction (important for polar materials), include
\texttt{DFPT} in your task list so Born effective charges are computed.
\end{notebox}

% ============================================================
\section{LOBSTER Bonding Analysis (COHP / COBI / COOP)}
\label{sec:lobster}
% ============================================================

Adding \texttt{LOBSTER COHP/COBI analysis} (keywords: \texttt{lobster},
\texttt{cohp}, \texttt{cobi}, \texttt{coop}, \texttt{bonding}) to the
task list creates \texttt{08\_lobster}: a symmetry-off NSCF that reads
\texttt{02\_scf/CHGCAR}, followed by a run of the \texttt{lobster}
binary.

\begin{itemize}
  \item \textbf{ISYM=0 is required} --- a symmetrised k-mesh gives
        unusable COHP/ICOHP values (wrong sign, errors above 100\%).
        The generated INCAR sets it automatically.
        With SOC enabled, \textbf{ISYM=$-$1} is used instead: spin-orbit
        coupling breaks time-reversal symmetry, which ISYM=0 would still
        assume.
  \item \texttt{NBANDS} is auto-set to at least the LOBSTER
        basis-function count; \texttt{LMAXMIX} is chosen from the
        elements (6 for $f$, 4 for $d$, 2 for $sp$).
  \item The LOBSTER binary is set with \texttt{LOBSTER\_X} in
        \texttt{site.env} (default \texttt{lobster-5.1.0-OSX}); it can
        be overridden at runtime via \texttt{\$LOBSTER\_BIN}.
\end{itemize}

\begin{lstlisting}
Project: NaCl
Methods: PBE functional
Tasks: SCF, LOBSTER COHP/COBI analysis
MPI: 16
\end{lstlisting}

Run with \texttt{./vasp-agent.py} (builds and runs \texttt{02\_scf} +
\texttt{08\_lobster}); the GUI shows COHP/COBI/COOP plots on the
Results tab.  To aggregate averaged COHP/COBI over many projects into
\texttt{lobster\_acohp\_acobi.csv}, run
\texttt{python3 lobster\_postprocess.py}.

% ============================================================
\section{Worked Examples}
% ============================================================

\subsection{MoS\textsubscript{2} monolayer with SOC}

\begin{lstlisting}
Project: MoS2 monolayer

Methods: PBEsol functional
         SOC with magnetization in z-direction

Tasks: structure relaxation
       SCF calculation
       band structure
       DOS (projected on Mo-d and S-p)

Convergence: Test k-points 8x8 to 20x20
             ENCUT 400-600 eV

MPI: 64
\end{lstlisting}

The generator detects hexagonal (hP) and uses G-M-K-G-A-L-H-A.

\subsection{Nickel: magnetism + GGA+U}

\begin{lstlisting}
Project: Nickel GGA+U

Methods: PBEsol functional
         collinear magnetization
         GGA+U with U=4.0 on Ni-d orbitals

Tasks: structure relaxation
       SCF calculation
       band structure
       DOS (projected on Ni-d and Ni-s)

MAGMOM: 2.0
MPI: 128
NODES: 2
NTASKS_PER_NODE: 64
WALLTIME: 08:00:00
\end{lstlisting}

With a primitive FCC POSCAR (angles $\approx 60^\circ$) the generator
detects cF and uses G-X-W-K-G-L-U-W.  Band and DOS plots show both spin
channels.

\subsection{GaAs: phonons with NAC}
\label{sec:example-gaas-nac}

\begin{lstlisting}
Project: GaAs phonons

Methods: PBEsol functional

Tasks: SCF calculation
       DFPT Born charges and dielectric
       phonons lattice dynamics

PHONONS_DIM: 4 4 4
PHONONS_MESH: 20 20 20
PHONONS_BAND: G X W L G K

MPI: 64     WALLTIME: 24:00:00
\end{lstlisting}

\subsection{High-pressure relaxation with custom INCAR tags}
\label{sec:ex-pressure}

A constant-pressure cell relaxation at 10~GPa, combined with raw INCAR
passthrough to tighten convergence and add tags the parser has no
dedicated keyword for:

\begin{lstlisting}
Project: Silicon 10GPa

Methods: PBEsol functional

Tasks: structure relaxation
       SCF calculation
       band structure

PRESSURE: 10 GPa          # constant-P relax -> ISIF=3, IBRION=2, PSTRESS=100

# Tags applied to every step
INCAR:
   LREAL  = .FALSE.
   ADDGRID = .TRUE.
   LMAXMIX = 4
END_INCAR

# Tighter electronic convergence for the relaxation only
INCAR relax:
   EDIFF  = 1E-8
   EDIFFG = -0.005
END_INCAR

ENCUT: 600
MPI: 64
\end{lstlisting}

The \texttt{01\_relax/INCAR} is generated with \texttt{ISIF=3},
\texttt{PSTRESS=100}, the global tags appended, and \texttt{EDIFF}/%
\texttt{EDIFFG} overridden to the per-step values; \texttt{02\_scf} and
\texttt{03\_bands} receive only the global block.

\subsection{Bi\textsubscript{2}Se\textsubscript{3}: topological insulator}

\begin{lstlisting}
Project: Bi2Se3 topological

Methods: PBEsol functional
         SOC with magnetization in z-direction

Tasks: structure relaxation
       SCF calculation
       band structure
       DOS (projected on Bi-p and Se-p)
       Wannier90 wannierization

Convergence: Test k-points from 8x8x4 to 16x16x8
             ENCUT 400-600 eV

WANNIER_NUM_WANN: 18
WANNIER_PROJ: Bi:p;Se:p
WANNIER_EWIN: -2.0 to 2.0
MPI: 128
\end{lstlisting}

% ============================================================
\section{Common OUTCAR Checks}
% ============================================================

\begin{lstlisting}[language=bash]
grep "reached required accuracy"   OUTCAR          # converged?
grep "energy  without"             OUTCAR | tail -1 # total energy
grep "E-fermi"                     OUTCAR | tail -1 # Fermi level
grep "magnetization (x)"           OUTCAR | tail -5 # moment
grep "TOTAL-FORCE"                 OUTCAR | tail -20
grep -A3 "BORN EFFECTIVE CHARGES"  OUTCAR | head -20
\end{lstlisting}

% ============================================================
\section{Troubleshooting}
% ============================================================

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Symptom & Solution \\ \midrule
POTCAR not found &
  Check \texttt{VASP\_POTCAR\_DIR} in \texttt{site.env} or GUI System
  Setup.  Use the POTCAR chooser for unusual variants
  (\texttt{Ga\_d}, \texttt{Fe\_sv}). \\[3pt]
\texttt{vasp\_std} not found &
  Set the full binary path in \texttt{site.env} or GUI System Setup. \\[3pt]
WARNING: may not have converged &
  Increase NSW (ionic) or NELM (electronic) in INCAR and re-run. \\[3pt]
Wrong bands ($W = \Gamma$) &
  Conventional cubic cell detected; use the \textit{primitive} cell
  for FCC/BCC so auto-detection picks cF/cI. \\[3pt]
Projected DOS missing &
  Check instructions.txt: \texttt{projected on} and \texttt{projections on}
  are both accepted.  Verify element names match POSCAR. \\[3pt]
KPAR error &
  KPAR $\times$ NCORE must divide total MPI tasks.  Use auto-defaults or
  set explicitly valid values. \\[3pt]
GUI unreachable &
  Check SSH tunnel: \texttt{ssh -L 5001:localhost:5001 user@cluster}. \\[3pt]
Project missing from dropdown &
  Folder must contain \texttt{POSCAR} + \texttt{settings.json} in the
  directory where \texttt{vasp-gui.py} was started. \\[3pt]
Band plot empty or not shown &
  Usually \texttt{sumo-bandplot} crashed on this band structure; the GUI now
  falls back to a built-in plotter automatically.  An empty plot also means a
  wrong $E_F$: the Fermi level is read from \texttt{04\_dos/OUTCAR} (then
  \texttt{02\_scf/OUTCAR}), since \texttt{03\_bands} eigenvalues share the
  NSCF/DOS energy reference. \\
\bottomrule
\end{tabularx}

% ============================================================
\section{Version History}
% ============================================================

\begin{description}
  \item[\textcolor{mainblue}{v1.0} (2026-03-23)]
    Initial release --- GUI, workstation agent, PBEsol/R2SCAN/HSE06,
    SOC, GGA+U, convergence tests, Wannier90, phonopy.
  \item[\textcolor{mainblue}{v1.1} (2026-04--05)]
    Per-task KPAR/NCORE smart defaults; SLURM agent with dependency-chained
    submission; cumulative DOS by element (no pymatgen); copy-from-relax/scf
    scripts.
  \item[\textcolor{mainblue}{v1.2} (2026-06)]
    Collinear magnetism: MAGMOM per atom, spin-resolved band and DOS plots.
    Auto k-path detection (8 crystal systems).
    Fixed DOS projection parsing.
    Separate GUI and agent quick-reference sheets.
    Rewritten comprehensive LaTeX manual.
  \item[\textcolor{mainblue}{v1.3} (2026-06-15)]
    Raw INCAR passthrough (\texttt{INCAR: \ldots\ END\_INCAR}, global or
    per-step) for arbitrary VASP tags.
    Constant-pressure relaxation via \texttt{PRESSURE} (ISIF=3, IBRION=2,
    PSTRESS).
  \item[\textcolor{mainblue}{v1.4} (2026-06-19)]
    High-throughput driver \texttt{ht-mp-scf.py}: Materials Project primitive
    cells $\to$ batch SCF + ELF, with per-material element/$Z$ table and
    cross-material \texttt{afterok} chaining on SLURM.
  \item[\textcolor{mainblue}{v1.5} (2026-06-24)]
    Robust band plotting across the GUI and both CLI agents (workstation +
    SLURM): the built-in \texttt{vasprun.xml} plotter now handles non-spin as
    well as spin, and is used automatically when \texttt{sumo-bandplot} crashes
    or is unavailable.  Fixed the matplotlib fallback (\texttt{plot\_results.py})
    to resolve the real Fermi level and emit PNG, not just PDF.
  \item[\textcolor{mainblue}{v1.6} (2026-07-08)]
    Band k-path derived from the real space group via spglib
    (Setyawan--Curtarolo) in all agents and the GUI, with the geometric
    classifier kept as fallback.
    New \texttt{08\_lobster} step: LOBSTER COHP/COBI/COOP bonding
    analysis with GUI plots and the \texttt{lobster\_postprocess.py}
    aggregator; the NSCF uses ISYM=0, or ISYM=$-$1 when SOC is on.
    Full GPL-3.0 text in \texttt{LICENSE} (GitHub now detects the
    license correctly).
    Manual and quick-reference sheets brought up to date.
  \item[\textcolor{mainblue}{v1.7} (2026-07-08)]
    DFPT step overhaul: KPAR/NCORE forced to 1 (linear response and
    PEAD support neither); automatic finite-field route
    (\texttt{LCALCEPS}, IBRION=6) when SOC, HSE06 or R2SCAN is
    requested; \texttt{extract\_born.py} now reports
    $\varepsilon_{\infty}$, the ionic contribution, total static
    $\varepsilon_{0}$, an acoustic-sum-rule check, element-labelled
    $Z^{*}$, and clamped-/relaxed-ion piezoelectric tensors.
  \item[\textcolor{mainblue}{v1.8} (2026-07-09)]
    ELF bond profiles: \texttt{elf\_bonds.py} plots the electron localization
    function along non-equivalent nearest-neighbour bonds (atom centre to atom
    centre), wired into \texttt{analyze.sh} for both CLI agents and shown in the
    GUI Results tab whenever \texttt{02\_scf/ELFCAR} exists.  Output is named
    after the reduced formula (\texttt{MoSe2\_elf\_bonds.pdf}).  Detects the
    all-zero \texttt{ELFCAR} that VASP writes for non-collinear / spin-orbit
    runs and reports it instead of emitting a blank plot.
  \item[\textcolor{mainblue}{v1.9} (2026-07-22)]
    Physics-based cell-aware k-mesh generation replacing fixed uniform grids.
    Two density tiers: \textbf{coarse} (1000 k-points per reciprocal atom,
    kpra) for convergence tests and relaxations; \textbf{fine} (5000~kpra) for
    SCF, DOS, and spectroscopic steps.  Subdivisions $N_i \propto |\mathbf{b}_i^*|$
    from reciprocal lattice vectors, all-even to include BZ boundary points;
    $N_z = 1$ for 2-D slabs.  Hexagonal systems: fixed $6{\times}6$ (coarse)
    or $12{\times}12$ (fine) in-plane, $N_z$ from $|\mathbf{b}_3^*|/|\mathbf{b}_1^*|$.
    DOS mesh always $2N_x{\times}2N_y{\times}2N_z$.  DFPT forced to match SCF
    mesh exactly.  New instruction key \texttt{KMESH\_DENSITY} (coarse/fine).
    GGA+U anion gate extended to all chalcogenides and halides
    (O, F, S, Se, Te, Cl, Br, I); Hubbard lookup table now covers sulfides
    such as AgCrP$_2$S$_6$.
\end{description}

% ============================================================
\section*{Addendum: High-Throughput Quick Reference}
\addcontentsline{toc}{section}{Addendum: High-Throughput Quick Reference}
\label{sec:addendum-ht}
% ============================================================

Cheat-sheet for \texttt{ht-mp-scf.py} (full description in
Section~\ref{sec:highthroughput}).  For each Materials Project ID it
downloads the \textbf{primitive} cell, stages an SCF calculation that also
writes the electron localization function (\texttt{LELF = .TRUE.}
$\rightarrow$ \texttt{ELFCAR}) on a cell-aware coarse $\Gamma$-mesh
(1000~kpra default; override with \texttt{KMESH\_DENSITY} or
\texttt{-{}-kmesh}), and builds a \texttt{runall.sh}.

\begin{lstlisting}[language=bash]
pip install pymatgen mp-api
export MP_API_KEY=...            # Materials Project key
export VASP_POTCAR_DIR=...       # POTCAR library (used by the agent)

printf 'mp-149\nmp-2534\n' > highthrouput_list   # one mp-id per line

./ht-mp-scf.py --agent local  --mpi 16           # workstation, sequential
./ht-mp-scf.py --agent slurm  --profile slurm    # cluster, afterok-chained
./ht-mp-scf.py --agent slurm  --no-chain         # cluster, independent

./runall.sh                      # build inputs + run/submit each SCF
\end{lstlisting}

\begin{tabularx}{\linewidth}{@{}lX@{}}
\thead
Output & Location \\ \midrule
ELFCAR (electron localization) & \texttt{<mp-id>/02\_scf/ELFCAR} \\
Staged inputs (POSCAR + instructions) & \texttt{\_ht\_inputs/<mp-id>/} \\
Element / $Z$ table & \texttt{\_ht\_inputs/elements\_Z.txt} \\
\bottomrule
\end{tabularx}

\begin{notebox}
\textbf{Transfer to a workstation:} only the project folder is needed
(code, \texttt{modules/}, \texttt{profiles/}); VASP binaries and the POTCAR
library already live there.  \texttt{ht-mp-scf.py} needs internet for the MP
download, so run it on the login node --- \texttt{runall.sh} does none.
\end{notebox}

\begin{notebox}
\textbf{If the download step errors with}
\texttt{'dict' object has no attribute 'number'}: that is an spglib /
pymatgen version mismatch.  Fix the root cause with
\texttt{pip install -U spglib}.  The driver also falls back to
\texttt{Structure.get\_primitive\_structure()} automatically, so a run is
never blocked by it.
\end{notebox}

\vfill
\noindent{\color{mainblue}\rule{\linewidth}{0.5pt}}\\[4pt]
\noindent\small
Source: \url{https://github.com/KeivanS/VASP-Flow}\hfill
Compiled: \today\hfill
Maintained by Keivan Esfarjani

\end{document}
